### 2024-03-23 R/analysis-code/exploratory-analysis.qmd
change dataset and finish part 3

---
title: "An example exploratory analysis script"
author: "Xueyan Hu"
date: "03/22/2024"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# fit simple linear model

First, I am curious that if physical activity could be correlated to the choice of transportation since I have heard a theory that if someone take much effort on commute transportation, they tend to spend less time on exercise.

```{r}
# Fit multinomial logistic regression model
#library(nnet)
#multinom_model <- multinom(FAF ~ MTRANS, data = mydata)

# Summary of the model
#summary(multinom_model)
```
I will try to figure it out later how I could make it fitting.

Then, I would like to see how drinking water plus physical activity affect obesity.

```{r}
# linear model of BMI by water and FAF
lm1 <- lm(BMI ~ Water + FAF + Water * FAF, data = mydata)
lm1_result <- summary(lm1)

# save the result
lm_file = here("results", "tables", "lm1table.rds")
saveRDS(lm1_result, file = lm_file)
```

In addition, I want to fit model for alcohol consumption and obesity level.

```{r}
mydata$Alcohol <- as.factor(mydata$Alcohol)
lm2 <- lm(BMI ~ Alcohol, data = mydata)
lm2_result <- summary(lm2)

# save the result
lm_file = here("results", "tables", "lm2table.rds")
saveRDS(lm2_result, file = lm_file)
```
I think it makes a little bit sense to me that there is a connection effect of water and physical activity on BMI, which might be supportive for what we learn that drinking more water and doing more exercise will lower the risk of obesity.

Last, I will try to use all predictors that I believe may affect BMI to fit a model.

```{r}
mydata1 <- mydata %>%
  select(History, Water, Alcohol, FAF, MTRANS, BMI)
lm3 <- lm(BMI ~ ., data = mydata1)
lm3_result <- summary(lm3)

# save the result
lm_file = here("results", "tables", "lm3table.rds")
saveRDS(lm3_result, file = lm_file)
```
# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.