### 2024-03-23 R/eda-code/eda.qmd

---
title: "Project eda"
author: "Xueyan Hu"
date: "2023-02-22"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# gender and age distribution for the cohort under investigation

```{r}
# Convert numeric values to labels
mydata$Gender <- ifelse(mydata$Gender == 1, "Male", "Female")

# bar chart of gender distribution
bc1 <- ggplot(mydata, aes(x = Gender)) +
  geom_bar(fill = c("pink", "skyblue"), color = "black", width = 0.5) +
  geom_text(stat = 'count', aes(label=..count..), vjust = -0.5, size=3, color="black") +
  labs(x = "Gender", y = "Count", title = "Bar Chart of Gender Distribution") +
  theme_minimal()
plot(bc1)

# histogram of population by age
hp2 <- ggplot(mydata, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "grey", color = "black") +
  labs(x = "Age", y = "Count", title = "Histogram of Age distribution") +
  stat_bin(binwidth = 5, geom = "text", aes(label = ..count..),
           vjust = -0.5, color = "black", size = 3)  +
  theme_minimal()
plot(hp2)
 # bar chart of obesity level distribution
bc3 <- ggplot(mydata, aes(x = Obesity)) +
  geom_bar() +
  geom_text(stat = 'count', aes(label=..count..), vjust = -0.5, size=3, color="black") +
  scale_x_discrete(labels = c("Under", "Normal", "OW1", "OW2", "Obese1", "Obese2", "Obese3")) +
  labs(x = "Obesity level", y = "Count", title = "Bar Chart of obesity Distribution") +
  theme_minimal()
plot(bc3)
```

# Explore a little bit deeper for Gender and Age

```{r}
unique(mydata$Obesity)
# obesity level by gender
sbp4 <- ggplot(mydata, aes(x = Obesity, fill = Gender)) +
  geom_bar(position = "stack") +
  labs(x = "Obesity Level", y = "Count", title = "Gender Distribution by Obesity Level") +
  scale_fill_manual(values = c("skyblue", "pink"), labels = c("Male", "Female")) +
  scale_x_discrete(labels = c("Under", "Normal", "OW1", "OW2", "Obese1", "Obese2", "Obese3")) +
  theme_minimal()
plot(sbp4)

# Obesity level by age
# Categorize age into groups
mydata <- mydata %>%
  mutate(Age_Group = cut(Age, breaks = c(0, 20, 30, 40, 50, 60, 70, Inf),
                         labels = c("0-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71+")))

sbp5 <- ggplot(mydata, aes(x = Obesity, fill = Age_Group)) +
  geom_bar(position = "stack",stat = "count") +
  coord_flip() + # Rotate the plot to make it horizontal
  labs(x = "Obesity level", y = "Count", title = "Stacked Bar Chart of Obesity Level by Age") +
  theme_minimal()
plot(sbp5)
```
I think the population distribution for age is not balanced since 20-30 age groups contains most of the observations so I will not explore the relationship between obesity and age.

# The relationship between family obesity history and obesity level

```{r}
vp6 <-ggplot(mydata, aes(x = History, y = BMI, fill = History)) +
  geom_violin() +  # Add violin geometry
  geom_boxplot(width = 0.1, fill = "white", color = "black") +  # Add boxplot for better visualization
  geom_jitter(aes(color = History), width = 0.2, alpha = 0.5, size = 1) +   # Add jittered scatterplot
  labs(x = "Family History of Obesity", y = "BMI", title = "Violin Plot of BMI by History of Obesity") +
  scale_fill_manual(values = c("yes" = "lightyellow", "no" = "lightgreen")) +  # Define colors
  theme_minimal()
plot(vp6)
```
# summary table

```{r}
summary1 <- summary(mydata)
summary2 <-skim(mydata)
summarytable_file = here("results", "tables", "summarytable1.rds")
saveRDS(summary1, file = summarytable_file)
summarytable_file = here("results", "tables", "summarytable2.rds")
saveRDS(summary2, file = summarytable_file)
```
# Saving all of the figures

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

```{r}
figure_file = here("results","figures","gender-distribution.png")
ggsave(filename = figure_file, plot=bc1) 
figure_file = here("results","figures","obesity-distribution.png")
ggsave(filename = figure_file, plot=bc3) 
figure_file = here("results","figures","obesity level by geder.png")
ggsave(filename = figure_file, plot=sbp4) 
figure_file = here("results","figures","obesity level by family history.png")
ggsave(filename = figure_file, plot=vp6) 
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.