### 2024-03-23 R/processing-code/processingfile-v1.qmd

---
title: "Project data cleaning"
author: "Xueyan Hu"
date: "2023-02-22"
output: html_document
---

# Processing script

This contains the same code and comments/information as `processingcode.R`.

This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.

See the other Quarto file for my currently preferred approach of pulling code from the R script into the Quarto file.

# Setup

### load packages

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

### loading data

```{r}
data_location <- here::here("data","raw-data","ObesityDataSet_raw_and_data_sinthetic.csv")
rawdata <- read.csv(data_location)
```

#### check on codebook

```{r}
codebook <- here::here("data","raw-data","codebook.xlsx")
read_excel(codebook)
```
Several ways of looking at the data

The meaning of each variable can be checked [online](https://www.kaggle.com/datasets/yasserh/titanic-dataset/data).

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

### data cleaning

First generating a new variable BMI using height and weight, as well as change some variable names and categorical levels.

```{r}
d1 <- rawdata %>%
  mutate(BMI = Weight / (Height)^2) %>%
  rename(History = family_history_with_overweight, Water = CH2O, Alcohol = CALC, Obesity = NObeyesdad) %>%
  mutate(Gender = if_else(Gender == "Male", 1, 2))

d2 <- d1 %>% 
  mutate(Alcohol = case_when(
    Alcohol == "no" ~ 1,
    Alcohol == "Sometimes" ~ 2,
    Alcohol == "Frequently" ~ 3,
    Alcohol == "Always" ~ 4,
    TRUE ~ NA_integer_  # Handle any other cases
  ))
# Print the updated dataframe
head(d2)
```

My interest variables are gender, age, alcohol consumption, water drinking, transportation, physical activity, and outcomes will be obesity level and BMI.

```{r}
d3 <- d2 %>%
  select(Gender, Age, History, Water, Alcohol, FAF, MTRANS,BMI, Obesity)
head(d3)
```

### Save data

Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. This preserves coding like factors, characters, numeric, etc. If you save as CSV, that information would get lost. However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data: http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
processeddata <- d3
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).