### 2024-04-04 R/analysis-code/exploratory-analysis.qmd
part 4

---
title: "An example exploratory analysis script"
author: "Xueyan Hu"
date: "03/22/2024"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# fit simple linear model

First, I am curious that if physical activity could be correlated to the choice of transportation since I have heard a theory that if someone take much effort on commute transportation, they tend to spend less time on exercise.

```{r}
# Fit multinomial logistic regression model
#library(nnet)
#multinom_model <- multinom(FAF ~ MTRANS, data = mydata)

# Summary of the model
#summary(multinom_model)
```
I will try to figure it out later how I could make it fitting.

Then, I would like to see how drinking water plus physical activity affect obesity.

```{r}
# linear model of BMI by water and FAF
lm1 <- lm(BMI ~ Water + FAF + Water * FAF, data = mydata)
lm1_result <- summary(lm1)
lm1_result

# save the result
lm_file = here("results", "tables", "lm1table.rds")
saveRDS(lm1_result, file = lm_file)
```

In addition, I want to fit model for alcohol consumption and obesity level.

```{r}
mydata$Alcohol <- as.factor(mydata$Alcohol)
lm2 <- lm(BMI ~ Alcohol, data = mydata)
lm2_result <- summary(lm2)

# save the result
lm_file = here("results", "tables", "lm2table.rds")
saveRDS(lm2_result, file = lm_file)
```
I think it makes a little bit sense to me that there is a connection effect of water and physical activity on BMI, which might be supportive for what we learn that drinking more water and doing more exercise will lower the risk of obesity.

I haven't include gender or age for model fitting, I will try to combine them in one model here.

```{r}
lm4 <- lm(BMI ~ Gender * Age, data = mydata)
lm4_result <- summary(lm4)
lm4_result

# save the result
lm_file = here("results", "tables", "lm3table.rds")
saveRDS(lm3_result, file = lm_file)
```

This linear model proves gender and age can also be predictors for BMI and there is interaction between them. So I will include them in the all predictor model.

Last, I will try to use all predictors that I believe may affect BMI to fit a model.

```{r}
lm3 <- lm(BMI ~ Gender + Age + History + Water + Alcohol + FAF + MTRANS, data = mydata)
lm3_result <- summary(lm3)
lm3_result

# save the result
lm_file = here("results", "tables", "lm3table.rds")
saveRDS(lm3_result, file = lm_file)
```

Choose predictors only important for the model fitting. Thanks to my classmates' replies, I received some useful suggestion for choosing predictors, which are chi-square, LASSO and lowest RMSE. 
According to my searching result online, chi-square is good for categorical variable as outcome, and LASSO might shrink the coefficients too much. So I will go with the lowest RMSE in the following and see what will happen next.

```{r}
# Load necessary packages
library(caret)

# Split data into training and testing sets
set.seed(123)  # Set seed for reproducibility
trainIndex <- createDataPartition(mydata$BMI, p = 0.8, list = FALSE)
trainData <- mydata[trainIndex, ]
testData <- mydata[-trainIndex, ]

# find important predictors without interactions

# Define predictors
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")

# Initialize variables to store results
minRMSE1 <- Inf
bestSubset1 <- NULL

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors)) {
  subsets <- combn(predictors, i)
  for (j in 1:ncol(subsets)) {
    predictors_subset <- subsets[, j]

    # Fit model using subset of predictors
    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
    model <- lm(formula, data = trainData)

    # Make predictions on test set
    predictions <- predict(model, newdata = testData)

    # Calculate RMSE
    RMSE1 <- sqrt(mean((testData$BMI - predictions)^2))

    # Update minRMSE and bestSubset if RMSE is lower
    if (RMSE1 < minRMSE1) {
      minRMSE1 <- RMSE1
      bestSubset1 <- predictors_subset
    }
  }
}

# Print results
print(paste("Best subset of predictors:", paste(bestSubset1, collapse = ", ")))
print(paste("Lowest RMSE:", minRMSE1))
```

```{r}
# Define predictors and response variable
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
interaction_terms <- c("Gender_Age", "FAF_Water")  # Define interaction terms
response <- "BMI"

# Create interaction terms in training and testing data
trainData$Gender_Age <- trainData$Gender * trainData$Age
trainData$FAF_Water <- trainData$FAF * trainData$Water

testData$Gender_Age <- testData$Gender * testData$Age
testData$FAF_Water <- testData$FAF * testData$Water

# Update list of predictors to include interaction terms
predictors_updated <- c(predictors, "Gender_Age", "FAF_Water")

# Initialize variables to store results
minRMSE2 <- Inf
bestSubset2 <- NULL

# Iterate through all possible combinations of predictors
for (i in 1:length(predictors_updated)) {
  subsets <- combn(predictors_updated, i)
  for (j in 1:ncol(subsets)) {
    predictors_subset <- subsets[, j]

    # Fit model using subset of predictors
    formula <- paste("BMI ~", paste(predictors_subset, collapse = " + "))
    model <- lm(formula, data = trainData)

    # Make predictions on test set
    predictions <- predict(model, newdata = testData)

    # Calculate RMSE
    RMSE2 <- sqrt(mean((testData$BMI - predictions)^2))

    # Update minRMSE and bestSubset if RMSE is lower
    if (RMSE2 < minRMSE2) {
      minRMSE2 <- RMSE2
      bestSubset2 <- predictors_subset
    }
  }
}

# Print results
print(paste("Best subset of predictors (including interactions):", paste(bestSubset2, collapse = ", ")))
print(paste("Lowest RMSE (including interactions):", minRMSE2))
```

It is under my expectation that all-predictor model with the 2 interactions has the lowest RMSE. So I also would like to try LASSO regression model here and see the result.

```{r}
# Load necessary packages
library(glmnet)
library(caret)

# Set seed for reproducibility
set.seed(123)

# Define predictors and response variable
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
response <- "BMI"

# Prepare data
x <- model.matrix(as.formula(paste(response, "~", paste(predictors, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
y <- mydata[, response]

# Fit LASSO regression model using cross-validation
lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression

# Extract optimal lambda value
best_lambda <- lasso_model$lambda.min

# Extract coefficients of predictors for the optimal lambda
coef <- coef(lasso_model, s = best_lambda)

# Identify non-zero coefficients (selected predictors)
selected_predictors1 <- rownames(coef)[coef != 0]

# Print selected predictors
print(paste("Selected predictors:", paste(selected_predictors1, collapse = ", ")))
```

```{r}
# Fit LASSO regression model using cross-validation with interaction terms
x <- model.matrix(as.formula(paste(response, "~", paste(predictors_updated, collapse = "+"))), data = mydata)[, -1]  # Exclude intercept column
y <- mydata[, response]
lasso_model <- cv.glmnet(x, y, alpha = 1, nfolds = 10)  # Use alpha = 1 for LASSO regression

# Extract optimal lambda value
best_lambda <- lasso_model$lambda.min

# Extract coefficients of predictors for the optimal lambda
coef <- as.matrix(coef(lasso_model, s = "lambda.min"))

# Identify non-zero coefficients (selected predictors)
selected_predictors <- rownames(coef)[coef != 0]

# Print selected predictors
print(paste("Selected predictors (including interactions):", paste(selected_predictors, collapse = ", ")))
```

The LASSO regression result includes some specific levels in variables instead of the entire variables.

Using CV to test evaluate the model with all predictors.

```{r}
# Define predictors and response variable
predictors <- c("Gender", "Age", "History", "Water", "Alcohol", "FAF", "MTRANS")
interaction_terms <- c("Gender_Age", "FAF_Water")
response <- "BMI"

# Define training control with repeated k-fold cross-validation
train_control <- trainControl(method = "repeatedcv", 
                              number = 10,       # Number of folds
                              repeats = 3)       # Number of repetitions

# Define the model formula
formula <- as.formula(paste(response, "~", paste(c(predictors, interaction_terms), collapse = " + ")))

# Train the model using cross-validation
model <- train(formula,
               data = mydata,
               method = "lm",                    # Use linear regression
               trControl = train_control,       # Use defined training control
               preProcess = c("center", "scale"),  # Preprocess data
               metric = "RMSE")                    # Use RMSE as the metric

# Print the cross-validation results
print(model)
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.