### 2024-04-22 products/manuscript/supplement/Supplementary-Material.qmd

---
title: "Supplement to Explore the factors that may be associated with obesity"

subtitle: ""
author: Xueyan Hu
date: "`r Sys.Date()`"
format:
  pdf:
    documentclass: scrartcl
    papersize: letter
    toc: false
    number-sections: true
    highlight-style: github
    bibliography: ../products/manuscript/dataanalysis-references.bib
csl: ../products/manuscript/apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

# Overview

All of the raw data file in csv, processed data in RDS file, descriptive analysis, exploratory analysis,and full data analysis codes and products are included in this project folder.

# Code and file information

-   assets: storing reference pdf files and citation style files

-   data:

    -   raw-data: codebook created by me and original data file downloaded from website

    -   processed-data: cleaned data saved in RDS file

-   products: manuscript in qmd file and its render output with same name in doc file

-   R: including all codes needed

    -   analysis-code: including codes for simple exploratory analysis and machine learning models fitting both in qmd file and R script file

    -   eda-code: including codes for simple exploratory analysis and machine learning models fitting both in qmd file and R script file

    -   processing-code: including codes for data cleaning process both in qmd and R script file

-   renv: storage for all of the packages used in this project in R

-   results:

    -   figures: all of the plots in png file created and saved from eda-code and analysis-code

    -   tables: summary tables created and saved from eda-code and 2 linear model tables from analysis-code

{{< pagebreak >}}

# Additional Method Details

I used linear regression model, LASSO model, random forest model for continuous variable BMI as the outcome; and I used discriminant analysis model and multinominal logistic regression for categorical variables obesity level as the outcome.

{{< pagebreak >}}

# Additional results

@tbl-summarytable1 shows a summary of the dataset according to each variable.

```{r}
#| label: tbl-summarytable1
#| tbl-cap: "a dataset summary"
#| echo: FALSE
summarytable1 = readRDS(here("results","tables","summarytable1.rds"))
knitr::kable(summarytable1)
```

@tbl-summarytable2 shows a more detailed summary of the dataset using skim function.

```{r}
#| label: tbl-summarytable2
#| tbl-cap: "a detailed dataset summary"
#| echo: FALSE
summarytable2 = readRDS(here("results","tables","summarytable2.rds"))
knitr::kable(summarytable2)
```

@tbl-lm1table shows linear model fitting result using Water + FAF + Water \* FAF as predictors and BMI as outcome.

```{r}
#| label: tbl-lm1table
#| tbl-cap: "linear model fit 1"
#| echo: FALSE
lm1table = readRDS(here("results","tables","lm1table.rds"))
# Extract coefficients from the summary object
coefficients <- coef(lm1table)

# Create a data frame from the coefficients
lm1table_df <- data.frame(
  Coefficients = rownames(coefficients),
  Estimate = coefficients[, 1],
  Std.Error = coefficients[, 2],
  t.value = coefficients[, 3],
  Pr..t.. = coefficients[, 4]
)

# Print the resulting data frame
knitr::kable(lm1table_df)
```

@tbl-lm2table shows linear model fitting result using Alcohol as predictors and BMI as outcome.

```{r}
#| label: tbl-lm2table
#| tbl-cap: "linear model fit 2"
#| echo: FALSE
lm2table = readRDS(here("results","tables","lm2table.rds"))
# Extract coefficients from the summary object
coefficients <- coef(lm2table)

# Create a data frame from the coefficients
lm2table_df <- data.frame(
  Coefficients = rownames(coefficients),
  Estimate = coefficients[, 1],
  Std.Error = coefficients[, 2],
  t.value = coefficients[, 3],
  Pr..t.. = coefficients[, 4]
)

# Print the resulting data frame
knitr::kable(lm2table_df)
```

@tbl-lm3table shows linear model fitting result using Gender + Age + History + Water + Alcohol + FAF + MTRANS as predictors and BMI as outcome.

```{r}
#| label: tbl-lm3table
#| tbl-cap: "linear model fit 3"
#| echo: FALSE
lm3table = readRDS(here("results","tables","lm3table.rds"))
# Extract coefficients from the summary object
coefficients <- coef(lm3table)

# Create a data frame from the coefficients
lm3table_df <- data.frame(
  Coefficients = rownames(coefficients),
  Estimate = coefficients[, 1],
  Std.Error = coefficients[, 2],
  t.value = coefficients[, 3],
  Pr..t.. = coefficients[, 4]
)

# Print the resulting data frame
knitr::kable(lm3table_df)
```

@tbl-lm4table shows linear model fitting result using Gender \* Age as predictors and BMI as outcome.

```{r}
#| label: tbl-lm4table
#| tbl-cap: "linear model fit 4"
#| echo: FALSE
lm4table = readRDS(here("results","tables","lm4table.rds"))
# Extract coefficients from the summary object
coefficients <- coef(lm4table)

# Create a data frame from the coefficients
lm4table_df <- data.frame(
  Coefficients = rownames(coefficients),
  Estimate = coefficients[, 1],
  Std.Error = coefficients[, 2],
  t.value = coefficients[, 3],
  Pr..t.. = coefficients[, 4]
)

# Print the resulting data frame
knitr::kable(lm4table_df)
```

{{< pagebreak >}}

# Discussion

For reproducing this project, raw-data folder will be downloaded and stored in a new folder called data and then included in a larger folder with all the codes together. Processing-code qmd file under R folder will be opened and run so the raw data will go through data cleaning process and the cleaned data file will be saved in a folder called process-data under the data folder. Then eda qmd file can be opened and run using processed data file and the outputs will be saved in a newly created fold called results. Results should include tables and figures two subfolders.

Then, exploratory-analysis qmd file under R folder will be involved for the whole data analysis and result will be seen directly in the file.

After all of the qmd files under R folder are run out, the final manuscript will be generated. There are reference related files under assets folders so it will be made sure download and then manuscript qmd file under products fold will be find and run with a doc file saved beside the qmd file.

Supplementary material file is under supplement subfolder and a pdf file will be created and saved after running this qmd file.

{{< pagebreak >}}

# References