### 2024-05-03 Rcode/dneda-code/dneda.qmd
improve after review

---
title: "Project eda"
author: "Xueyan Hu"
date: "2023-02-22"
output: html_document
---

This file loads the processed data cleaned from processingfile-v1 and basic descriptive analysis is applied in this file.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# gender and age distribution for the cohort under investigation

```{r}
# Convert numeric values to labels
mydata$Gender <- ifelse(mydata$Gender == 1, "Male", "Female")

# bar chart of gender distribution
bc1 <- ggplot(mydata, aes(x = Gender)) +
  geom_bar(fill = c("pink", "skyblue"), color = "black", width = 0.5) +
  geom_text(stat = 'count', aes(label=..count..), vjust = -0.5, size=3, color="black") +
  labs(x = "Gender", y = "Count", title = "Bar Chart of Gender Distribution") +
  theme_minimal()
plot(bc1)

# histogram of population by age
# Calculate the bin edges and labels
bin_data <- mydata %>%
  mutate(bin = cut(Age, breaks = c(0, 10, 20, 30, 40, 50, 60, 70, Inf), 
                   labels = FALSE, include.lowest = TRUE)) %>%
  group_by(bin) %>%
  summarise(bin_min = min(Age), bin_max = max(Age)) %>%
  mutate(bin_min = as.integer(bin_min),  # Convert bin_min to integer
         bin_max = as.integer(bin_max),  # Convert bin_max to integer
         bin_label = sprintf("%d-%d", bin_min, bin_max))  # Format bin labels

hp2 <- ggplot(mydata, aes(x = Age)) +
  geom_histogram(binwidth = 10, fill = "grey", color = "black") +
  labs(x = "Age", y = "Count", title = "Histogram of Age distribution") +
  stat_bin(binwidth = 10, geom = "text", aes(label = ..count..),
           vjust = -0.5, color = "black", size = 3)  +
  geom_text(data = bin_data, aes(x = bin_min + (bin_max - bin_min)/2 - 5, y = -5, label = bin_label),
            vjust = +1.2, size = 3, color = "black") +  # Add text labels for bins
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8))  # Adjust the size of x-axis text
plot(hp2)

 # bar chart of obesity level distribution
bc3 <- ggplot(mydata, aes(x = Obesity)) +
  geom_bar() +
  geom_text(stat = 'count', aes(label=..count..), vjust = -0.5, size=3, color="black") +
  scale_x_discrete(labels = c("Under", "Normal", "OW1", "OW2", "Obese1", "Obese2", "Obese3")) +
  labs(x = "Obesity level", y = "Count", title = "Bar Chart of obesity Distribution") +
  theme_minimal()
plot(bc3)
```

# Explore a little bit deeper for Gender and Age

```{r}
unique(mydata$Obesity)
# obesity level by gender
sbp4 <- ggplot(mydata, aes(x = Obesity, fill = Gender)) +
  geom_bar(position = "stack") +
  labs(x = "Obesity Level", y = "Count", title = "Gender Distribution by Obesity Level") +
  scale_fill_manual(values = c("skyblue", "pink"), labels = c("Male", "Female")) +
  scale_x_discrete(labels = c("Under", "Normal", "OW1", "OW2", "Obese1", "Obese2", "Obese3")) +
  theme_minimal()
plot(sbp4)

# Obesity level by age
# Categorize age into groups
mydata <- mydata %>%
  mutate(Age_Group = cut(Age, breaks = c(0, 20, 30, 40, 50, 60, 70, Inf),
                         labels = c("0-20", "21-30", "31-40", "41-50", "51-60", "61-70", "71+")))

sbp5 <- ggplot(mydata, aes(x = Obesity, fill = Age_Group)) +
  geom_bar(position = "stack",stat = "count") +
  coord_flip() + # Rotate the plot to make it horizontal
  labs(x = "Obesity level", y = "Count", title = "Stacked Bar Chart of Obesity Level by Age") +
  theme_minimal()
plot(sbp5)
```
I think the population distribution for age is not balanced since 20-30 age groups contains most of the observations so I will not explore the relationship between obesity and age.

# The relationship between family obesity history and obesity level

```{r}
vp6 <-ggplot(mydata, aes(x = History, y = BMI, fill = History)) +
  geom_violin() +  # Add violin geometry
  geom_boxplot(width = 0.1, fill = "white", color = "black") +  # Add boxplot for better visualization
  geom_jitter(aes(color = History), width = 0.2, alpha = 0.5, size = 1) +   # Add jittered scatterplot
  labs(x = "Family History of Obesity", y = "BMI", title = "Violin Plot of BMI by History of Obesity") +
  scale_fill_manual(values = c("yes" = "lightyellow", "no" = "lightgreen")) +  # Define colors
  theme_minimal()
plot(vp6)
```
# summary table

```{r}
summary1 <- summary(mydata)
summary2 <-skim(mydata)
summarytable_file = here("results", "tables", "summarytable1.rds")
saveRDS(summary1, file = summarytable_file)
summarytable_file = here("results", "tables", "summarytable2.rds")
saveRDS(summary2, file = summarytable_file)
```
# Saving all of the figures

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

```{r}
figure_file = here("results","figures","gender-distribution.png")
ggsave(filename = figure_file, plot=bc1) 
figure_file = here("results","figures","age-distribution.png")
ggsave(filename = figure_file, plot=hp2) 
figure_file = here("results","figures","obesity-distribution.png")
ggsave(filename = figure_file, plot=bc3) 
figure_file = here("results","figures","obesity level by geder.png")
ggsave(filename = figure_file, plot=sbp4) 
figure_file = here("results","figures","obesity level by age.png")
ggsave(filename = figure_file, plot=sbp5)
figure_file = here("results","figures","obesity level by family history.png")
ggsave(filename = figure_file, plot=vp6) 
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.