### 2024-05-03 Rcode/mlmodel-code/ml-model-v2.qmd

---
title: "An example exploratory analysis script with code pulled in"
author: "Yao Lu, Andreas Handel"
date: "2023-03-13"
output: html_document
---

This Quarto file has the content of machine learning model analysis.

This is essentially the same as the other `ml-model-v1` Quarto file, but now the code is not inside this file. Instead, it is pulled in from the R script `ml-model.R` using the code chunk labels.

# Setup
Load the chunk.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('exploratorycode.R')
```

Load the packages.
```{r,packages, echo=FALSE,message=FALSE}
```

Load the data.

```{r,loaddata}
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r,logistic}
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through through fit model

## lm1

```{r,lm1}
```

## lm2

```{r,lm2}
```

## lm3

```{r,lm3}
```

## lm4

```{r,lm4}
```

## model training

```{r,fitting linear models}
```

```{r,fitting logistic models}
```

## model assessment using testing data

```{r,test models}
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.