### 2024-05-03 Rcode/processing-code/processingfile-v1.qmd

---
title: "Project data cleaning"
author: "Xueyan Hu"
date: "2023-02-22"
output: html_document
---

# Processing script

# Setup

### load packages

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

### loading data

```{r}
data_location <- here::here("data","raw-data","ObesityDataSet_raw_and_data_sinthetic.csv")
rawdata <- read.csv(data_location)
```

#### check on codebook

```{r}
codebook <- here::here("data","raw-data","codebook.xlsx")
read_excel(codebook)
```

Several ways of looking at the data

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

### data cleaning

First generating a new variable BMI using height and weight, as well as change some variable names and categorical levels. For more intuitive observation and writing, I will change gender categories to 1 and 2 and name alcohol consumption levels to from 1 to 4.

```{r}
d1 <- rawdata %>%
  mutate(BMI = Weight / (Height)^2) %>%
  rename(History = family_history_with_overweight, Water = CH2O, Alcohol = CALC, Obesity = NObeyesdad) %>%
  mutate(Gender = if_else(Gender == "Male", 1, 2))

d2 <- d1 %>% 
  mutate(Alcohol = case_when(
    Alcohol == "no" ~ 1,
    Alcohol == "Sometimes" ~ 2,
    Alcohol == "Frequently" ~ 3,
    Alcohol == "Always" ~ 4,
    TRUE ~ NA_integer_  # Handle any other cases
  ))
# Print the updated dataframe
head(d2)
```

My interest variables are gender, age, alcohol consumption, water drinking, transportation, physical activity, and outcomes will be obesity level and BMI.

```{r}
d3 <- d2 %>%
  select(Gender, Age, History, Water, Alcohol, FAF, MTRANS,BMI, Obesity)
head(d3)
```

### Save data

```{r}
processeddata <- d3
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```